Relationships in the extremes and their influence on competition and coexistence

### 2022-03-07 paper/Paper.Rnw
readding changes intro.p2

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between two variables provide only a small 
portion of the overall information content of the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in Fig. \ref{fig:noise} all have the same
Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
of association are also often used, in addition to correlations
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
The variables in Fig. \ref{fig:noise}a,c 
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
and related ideas had previously been seldom 
considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.

%Intro to storage effects and why they might be influenced by ATAs
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance). 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
in times when the environment is poor. Dormant stages such as seed banks are an
example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species are currently rare.
Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.

>>>>

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between two variables provide only a small 
portion of the overall information content of the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in Fig. \ref{fig:noise} all have the same
Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
of association are also often used, in addition to correlations
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
The variables in Fig. \ref{fig:noise}a,c 
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
and related ideas had previously been seldom 
considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.

%Intro to storage effects and why they might be influenced by ATAs
It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance). 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
in times when the environment is poor. Dormant stages such as seed banks are an
example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species are currently rare.
Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.
